---
title: "L1 Overview"
author: "Xiaoting Chen"
date: "2023-08-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(HistData)
library(lubridate)
library(ggplot2)
library(gridExtra)
library(scales)
```

## Basic calculations

```{r}
# complex number
1i
round(exp(1i*pi)+1)
```

### practice 1.
```{r}
fib <- c(0,1,1,2,3,5,8,13,21,34) # 3rd number = sum(the two numbers before)

# alternatively 
fib = vector()
fib[1:2] = c(0,1)
for (i in 3:10) {
  fib[i] <- sum(c(fib[i-1],fib[i-2]))
}

length(fib)
sum(fib)
prod(fib)
diff(fib)

a <- 1:100
N <- 100
sum(a) == N*(N+1)/2

a_2 <- a^2
sum(a_2)
```

add up the first 100 integers without using sum() function
```{r}
n <- 100
x <- 1:n
s <- 0
for (i in 1:n) {
  s <- s + x[i]
}
```

add only even numbers in 1:100
```{r}
n <- 100
x <- 1:n
s <- 0
for (i in 1:n) {
  if (x[i]%%2 == 0) {
    s <- s + x[i] 
  }
}
```

## R basics

### list

```{r}
list_x <- list(A = pi,
               B = c(0,1),
               c = 1:10,
               D = c("one","two"))
list_x[2]
list_x[[2]]
```

### data frame

```{r}
data("Virginis.interp")
# Virginis.interp |> dplyr::as_tibble() |> class()
#virginis <- as_tibble(Virginis.interp)
virginis = Virginis.interp |> as_tibble()
mean(virginis$velocity)
```


### plots

```{r}
data("Snow.dates")
clr <- ifelse(Snow.dates$date < mdy("09081854"), "red", "darkgreen")
plot(deaths ~ date, data = Snow.dates,
     type = "h", lwd = 2, col = clr, xlab="")
points(deaths ~ date, data = Snow.dates,
       cex = 0.5, pch = 16, col = clr)
```

Intro to ggplot
```{r}
x <- seq(0, 100, by = 5)
y <- x^2

quadratic <- tibble(x = x, y = y)
p1 <- ggplot(data = quadratic, 
             mapping = aes(x = x, y = y))
p2 <- p1 + geom_point(size = 0.5)
p3 <- p1 + geom_line(linewidth = 0.2, 
                     color = 'red')
p4 <- p1 + geom_point(size = 0.5) + 
  geom_line(linewidth = 0.2, color = 'red')

grid.arrange(p1, p2, p3, p4, nrow = 2)
```


## simulating coin flips

```{r}
coin <- c("H", "T")
set.seed(0)
sp <- sample(coin, 100, replace = T, prob = c(0.6, 0.4)) 
mean(sp == "H")
```

check convergence of coin flips

```{r}
coin <- c(0,1)
set.seed(1) # why do we do this?
n <- 1e4
est_prop <- numeric(n) # allocate a vector of size n
for (i in 1:n) {
  x <- sample(coin, i, replace = TRUE)
  est_prop[i] <- mean(x)
}

coin_dat <- tibble(n = 1:n, est_prop = est_prop)
ggplot(data = coin_dat, mapping = aes(x = n, y = est_prop)) +
  geom_line(size=0.1)
```

Law of large numbers
```{r}
ggplot(data = coin_dat, mapping = aes(x = n, y = est_prop)) +
  geom_line(size = 0.1) +
  geom_hline(yintercept = 0.5, size = 0.2, color = 'red') +
  scale_x_continuous(trans = 'log10', label = comma) +
  xlab("Number of flips on Log10 scale") +
  ylab("Estimated proportion of Heads") +
  ggtitle("Error decreases with the size of the sample")
```


## Reproduce experiments / simulations
```{r}
estimate_proportion <- function(n) {
  coin <- c(1, 0)
  x <- sample(coin, n, replace = TRUE)
  est <- mean(x)
  return(est)
}
x <- replicate(1e3, estimate_proportion(10))
hist(x, 
     xlab = "Number of simulations", 
     main = "Distribution of Proportions of Heads")
```

```{r}
library(purrr)
par(mar = c(3, 3, 2, 1), mgp = c(2, .7, 0), tck = -.01, bg = "#f0f1eb")
y <- map_dbl(2:500, estimate_proportion). # the same as sapply(2:500, estimate_proportion)
plot(2:500, y, xlab = "", ylab = "", type = 'l')
```

practice: \
modify the coin bias to 0.3
```{r}
estimate_proportion_bias <- function(n, bias = 0.5) {
  coin <- c(1, 0)
  prob <- c(bias, 1-bias)
  x <- sample(coin, n, prob = prob, replace = TRUE)
  est <- mean(x)
  return(est)
}
x <- replicate(1e3, estimate_proportion_bias(n=10, bias = 0.3))
hist(x, 
     xlab = "Number of simulations", 
     main = "Distribution of Proportions of Heads")
```



## Estimating \(\pi\) by Simulation 
$$ \pi = 4\frac{A_c}{A_s}$$
(deviration in slides)
```{r}
set.seed(0)
n <- 1e3
x <- runif(n, -1, 1); y <- runif(n, -1, 1)
inside <- x^2 + y^2 < 1
data <- tibble(x, y, inside)
ggplot(aes(x = x, y = y), data = data)+ 
  geom_point(aes(color = inside)) + 
  theme(legend.position = "none") +
  coord_fixed(ratio = 1)
```

```{r}
cat("Estimated value of pi =", 4*sum(inside) / n)
```

practice:\
function to estimate value of pi with specified number of iterations
```{r}
set.seed(0)
est_pi <- function(n) {
  n <- n
  x <- runif(n, -1, 1); y <- runif(n, -1, 1)
  inside <- x^2 + y^2 < 1
  return(4*sum(inside) / n)
}

par(mar = c(3, 3, 2, 1), mgp = c(2, .7, 0), tck = -.01, bg = "#f0f1eb")
y <- map_dbl(2:1000, est_pi)
plot(2:1000, y, xlab = "", ylab = "", type = 'l')
abline(h = pi, col = "red", lty = 2)
```